### 2017-09-19 Paper_Notes.Rmd
Initial Commit

---
title: "Paper_Notes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Watson & Crick, 1953

- Watson, J. D., and F. H. Crick. "Molecular Structure of Nucleic Acids;
a Structure for Deoxyribose Nucleic Acid." Nature 171, no. 4356 (April 25,
1953): 737-38

**Summary:**

In this landmark paper, Watson and Crick propose a novel structure for deoxyribonucleic acid (DNA). Prior to their work, others structures were proposed including a model consisting of three intertwined chains, with phosphate groups located internally near the fiber axis and bases on the outside (Pauling and Corey, 1953). It is argued that this structure cannot work because negatively charged phosphate groups in close proximity would repel each other. Furthermore, the van der Waals forces generated by this model would be too small to hold it together. Instead, Watson and Crick propose a model consisting of two helical chains, coiled around the same axis with bases on the inside and phosphates on the outside. Although this structure is based on a model proposed by Furberg (1952), Watson and Crick's novelty stems from the manner in which the two helices are held together. A single base from one chain is bound to a single base from the opposite strand via hydrogen bonds. In order for this to work, one of the pair must have an underlying purine structure while the other must have an underlying pyrimidine structure. Assuming the bases only occur in the structure in the most plausible tautomeric form, only specific pairs of bases can bond together (i.e. adenine (purine) with thymine (pyrimidine), and guanine (purine) with cytosine (pyrimidine). I am both surprised and amazed by this paper. It is incredible that based only on certain assumptions, Watson and Crick were able to deduce the structure of DNA with exact dimensions. However, I am surprised that it is so short. Although they do say full details will be published elsewhere, it is hard to imagine editorials proposing radical shifts in certain scientific paradigms could be published without supporting evidence, data, or detailed calculations.